Unsupervised Detection and Tracking of Multiple Objects with Dependent Dirichlet Process Mixtures

### 2012-01-08 ltx/paper.tex
a few quick edits on inference section

A value proportional to the posterior probability can be computed for each possible value that $c_{i,t}$ may take on; these values allow us to construct a discrete probability distribution that may be drawn from to get a sample from the posterior distribution over assignments. The possible values that the assignment may take on are $c_{i,t} = k \in \{ 1 , \ldots ,  K_{t}^{'}, k^{*}\}$, where $K_{t}^{'}$ is the number clusters with a non-zero size at any point between time $t$ and the assignment's associated deletion time $d_{i,t}$, and $k^{*}$ is a potential new cluster. The probability that $c_{i,t}$ is assigned to a cluster $k$ can be computed for each $k \in \{ 1 , \ldots ,  K_{t}^{'}, k^{*} \}$, given all other members of the state, by
\begin{equation}
\begin{split}
p(c_{i,t} = k | \text{State}) & \propto
\prod_{i'=1}^{N_{t}}  \text{CRP}(c_{i',t} | \bold{m}_{t-1}, \alpha) \\
& \times \prod_{t' = t+1}^{d_{i,t}}  \prod_{i'=1}^{N_{t'}}   \text{CRP}(c_{i',t'} | \bold{m}_{t'-1}, \alpha) \\
 & \times
\begin{cases}
	P(\bold{x}_{i,t} | \theta_{k,t})                                        \hspace{12pt} \text{if} \hspace{4pt} k \in \{ 1, \ldots, K_{t}^{'} \} \\
	\int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta) d\theta    \hspace{12pt} \text{if}  \hspace{4pt}  k = k^{*}
\end{cases}
\end{split}
\end{equation}
% where $t$ is the time of observation $i$, and the distribution under the product $P(c_{i'} | m_{1:K_{t}^{'}, i'-1}, c_{1:i'-1})$ is given by 
%\begin{equation}
%P(c_{i} = k | m_{1:K_{t}^{'}, i-1}, c_{1:i-1}, \alpha) \propto 
%\begin{cases}
%m_{k, i-1} \hspace{12pt} \text{if} \hspace{4pt} k \in \{ 1, \ldots, K_{t}^{'} \} \\
%\alpha \hspace{12pt} \text{if}  \hspace{4pt}  k = k^{*}
%\end{cases}
%\end{equation}
where the distribution CRP is given by \eqref{crp_rep}, and the sizes of the clusters $\bold{m}_{t'-1}$ are calculated given the fact that $c_{i,t} = k$. Note that the above integral can be determined analytically to give
\begin{equation}
\label{marginal_obs_prob}
\int P(\bold{x}_{i} | \theta) \mathbb{G}_{0}(\theta)d\theta = \text{Student}_{\nu_{n} - 1}(\mu_{n}, \Lambda_{n}/(\kappa_{n}(\nu_{n}-1)))
\end{equation}
where ``Student'' denotes the Student's t distribution, where we follow the three-value parameterization (location parameter, scale parameter, and degrees of freedom) given in \cite{gelman2004bayesian}.

@@

\subsubsection{Sampling Auxiliary Variables $z_{k,t,m}$}
The posterior distribution of each of the auxiliary variables $z_{k,t,m}$ can be written as
\begin{equation}
P(z_{k,t,m} | \text{State}) \propto  P(z_{k,t,m} | \theta_{k,t-1}) P(\theta_{k,t} | z_{k,t,1:M})
\end{equation}
This distribution does not allow for samples to be easily drawn and thus requires a more sophisticated MCMC technique. In particular, the Metropolis-Hastings (M-H) algorithm is used to gain samples from the posterior distribution for the auxiliary variables; it involves constructing a markov chain with the sought posterior as its stationary distribution, and sampling sequentially from this constructed chain to get an approximate posterior sample. The M-H algorithm is outlined in Algorithm 1.%Algorithm~\ref{alg:MH}. \willie{am i citing wrong?}

\begin{algorithm}[h!]
\label{alg:MH}
\caption{Metropolis Hastings Algorithm}
\begin{algorithmic}[1]
\STATE $f(x)$ is the desired stationary PDF
\STATE $P(x_{i} | x_{i-1})$ is the proposal pdf
\STATE Initialize $x_{0}$
\FOR{$i = 1 : burn\hspace{1pt}in + N$}
\STATE Draw $x^{*} \sim P(x_{i} | x_{i-1})$
\STATE Draw $\alpha \sim U(0,1)$
\STATE $r_{accept} \leftarrow \frac{f(x^{*})P(x_{i-1} | x^{*})} {f(x_{i-1})P(x^{*} | x_{i-1})} $
\IF{$\alpha < min \{ 1, r_{accept} \} $}
\STATE $x_{i} \leftarrow x^{*}$
\ELSE
\STATE $x_{i} \leftarrow x_{i-1}$
\ENDIF
\ENDFOR
\STATE Discard samples $x_{1}, \ldots, x_{burn\hspace{1pt}in}$
\STATE \textbf{Output}: Remaining samples $x_{1}, \ldots, x_{N}$
\end{algorithmic}
%\caption{Say thing at end}
\end{algorithm}

The M-H algorithm is used to sample auxiliary variables for each alive cluster. In short, we sample a new value for all $M$ auxiliary variables, denoted $z_{k,t,m}^{*}$, according to the following proposal distribution
\begin{equation}
z_{k,t,m}^{*}  \sim  P(z_{k,t,m} | \theta_{k,t}) = \mathcal{N}(z_{k,t,m}^{s} | \boldsymbol{\mu}_{k,t}, \Sigma_{k,t}) \mathcal{M}n(z_{k,t,m}^{c} | \bold{p}_{k,t})
\end{equation}
compute the standard M-H acceptance ratio, which in this case (due to the proposal distribution choice) simplifies to 
\begin{equation}
r_{accept} = \frac{P(\theta_{k,t-1} | z_{k,t,m}^{*})}{P(\theta_{k,t-1} | z_{k,t,m})}
\end{equation}
and keep this sample with probability $min(1, r_{accept})$.

@@

This section details a Sequential Monte Carlo (SMC) sampler---also known as a particle filter---used to perform inference, which provides another method to carry out unsupervised detection and tracking. The SMC inference algorithm is given in Algorithm~\ref{alg:SMC}. In this algorithm, at each time step $t \in \{ 1, \ldots, T \}$, a number of samples refered to as ``particles'' are generated; each particle consists of a sample from the posterior distribution over the assignment for each observation, $c_{1,t}, \ldots, c_{N_{t}, t}$, parameters for each cluster $\theta_{1,t}, \ldots, \theta_{K_{t}, t}$, and size after deletion for each cluster $m_{1,t}, \ldots, m_{K_{t},t}$. A set of particles is sampled at each time step from relevant proposal distributions (described in Sections~\ref{sec:smc_proposal_1}, \ref{sec:smc_proposal_2}, and \ref{sec:smc_proposal_3}), a weight is computed for each particle, and a subset of the particles are resampled and kept via a resampling process. Furthermore, within each time step, Gibbs sampling is used to generate the samples associated with each particle.

The sequence of target distributions for the SMC algorithm may be written as
\begin{equation}
\begin{split}
%\pi_{t}(c_{1:N_{1}, 1}, \ldots, c_{1:N_{t}, t}, & \theta_{1:K_{1}, 1}, \ldots, \theta_{1:K_{t}, t}, m_{1:K_{1}, 1}, \ldots, m_{1:K_{t}, t}) = \\
%\pi_{t-1}(c_{1:N_{1}, 1}, \ldots, c_{1:N_{t-1}, t-1}, & \theta_{1:K_{1}, 1}, \ldots, \theta_{1:K_{t-1}, t-1}, m_{1:K_{1}, 1}, \ldots, m_{1:K_{t-1}, t-1}) \\
\pi_{t}(\bold{c}_{1:t}, \Theta_{1:t}, \bold{m}_{1:t+1}) & =  \pi_{t-1}(\bold{c}_{1:t-1}, \Theta_{1:t-1}, \bold{m}_{1:t})\\
& \times \prod_{i=1}^{N_{t}} P(c_{i,t} | \bold{m}_{t}, \Theta_{t}, \bold{c}_{1:t}, \bold{x}_{1:N_{t}}) \\
& \times \prod_{k=1}^{K} 
\begin{cases}
\begin{split}
P(\theta_{k,t} | \theta_{k,t-1}) \hspace{2mm} & \text{if} \hspace{2mm} k \leq K_{t-1} \\
\mathbb{G}_{0} \hspace{18mm} & \text{if} \hspace{2mm} k > K_{t-1}
\end{split}
\end{cases} \\
& \times \prod_{k=1}^{K} \text{DEL}(m_{k,t+1} | m_{k,t}, c_{1:N_{t-1}, t-1}, \rho)
\end{split}
\end{equation}
where $\bold{c}_{t} = c_{1,t}, \ldots, c_{N_{t}, t}$, $\Theta_{t} = \theta_{1,t}, \ldots, \theta_{K_{t}, t}$, and $\bold{m}_{t} = m_{1,t}, \ldots, m_{K_{t}, t}$ A few distributions referenced in the SMC algorithm are described in detail in the following sections.

@@

The probability of assignments given current cluster sizes, cluster parameters, and the Dirichlet Process granularity parameter $\alpha$ can be written as
\begin{equation}
\label{smc_assig_eqn}
\begin{split}
P \left( c_{i,t} | m_{1:K_{t}, t}, \theta_{1:K_{t}, t}, \alpha \right) &\propto\\
\text{CRP}(m_{1:K_{t}, t}, \alpha) &\times
\begin{cases}
P(\bold{x}_{i,t} | \theta_{c_{i,t},t}) \hspace{2mm} \text{if} \hspace{2mm} k \leq K_{t-1} \\
\int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta)d\theta \hspace{2mm} k > K_{t-1}
\end{cases}
\end{split}
\end{equation}
The integral can $\int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta)d\theta$ can be determined analytically, and is given in \eqref{marginal_obs_prob}.

@@

\section*{Discussion}
\label{sec:discussion}

@@

\section*{Conclusion}

>>>>

A value proportional to the posterior probability can be computed for each possible value that $c_{i,t}$ may take on; these values allow us to construct a discrete probability distribution that may be drawn from to get a sample from the posterior distribution over assignments. The possible values that the assignment may take on are $c_{i,t} = k \in \{ 1 , \ldots ,  K_{t}^{'}, k^{*}\}$, where $K_{t}^{'}$ is the number clusters with a non-zero size at any point between time $t$ and the assignment's associated deletion time $d_{i,t}$, and $k^{*}$ is a potential new cluster. The probability that $c_{i,t}$ is assigned to a cluster $k$ can be computed for each $k \in \{ 1 , \ldots ,  K_{t}^{'}, k^{*} \}$, given all other members of the state, by
\begin{equation}
\begin{split}
p(c_{i,t} = k | \text{State}) & \propto
\prod_{i'=i}^{N_{t}}  \text{CRP}(c_{i',t} | \bold{m}_{t-1}, \alpha) \\
& \times \prod_{t' = t+1}^{d_{i,t}}  \prod_{i'=1}^{N_{t'}}   \text{CRP}(c_{i',t'} | \bold{m}_{t'-1}, \alpha) \\
 & \times
\begin{cases}
	P(\bold{x}_{i,t} | \theta_{k,t})                                        \hspace{12pt} \text{if} \hspace{4pt} k \in \{ 1, \ldots, K_{t}^{'} \} \\
	\int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta) d\theta    \hspace{12pt} \text{if}  \hspace{4pt}  k = k^{*}
\end{cases}
\end{split}
\end{equation}
% where $t$ is the time of observation $i$, and the distribution under the product $P(c_{i'} | m_{1:K_{t}^{'}, i'-1}, c_{1:i'-1})$ is given by 
%\begin{equation}
%P(c_{i} = k | m_{1:K_{t}^{'}, i-1}, c_{1:i-1}, \alpha) \propto 
%\begin{cases}
%m_{k, i-1} \hspace{12pt} \text{if} \hspace{4pt} k \in \{ 1, \ldots, K_{t}^{'} \} \\
%\alpha \hspace{12pt} \text{if}  \hspace{4pt}  k = k^{*}
%\end{cases}
%\end{equation}
where the distribution CRP is given by \eqref{crp_rep}, and the sizes of the clusters $\bold{m}_{t'-1}$ are calculated given the fact that $c_{i,t} = k$. Note that the above integral can be determined analytically to give
\begin{equation}
\label{marginal_obs_prob}
\int P(\bold{x}_{i} | \theta) \mathbb{G}_{0}(\theta)d\theta = \text{Student}_{\nu_{n} - 1}(\mu_{n}, \Lambda_{n}/(\kappa_{n}(\nu_{n}-1)))
\end{equation}
where ``Student'' denotes the Student's t distribution, where we follow the three-value parameterization (location parameter, scale parameter, and degrees of freedom) given in \cite{gelman2004bayesian}.

@@

\subsubsection{Sampling Auxiliary Variables $z_{k,t,m}$}
The posterior distribution of each of the auxiliary variables $z_{k,t,m}$ can be written as
\begin{equation}
\label{stationary_pdf}
P(z_{k,t,m} | \text{State}) \propto  P(z_{k,t,m} | \theta_{k,t-1}) P(\theta_{k,t} | z_{k,t,1:M})
\end{equation}
This distribution does not allow for samples to be easily drawn and thus requires a more sophisticated MCMC technique. In particular, the Metropolis-Hastings (M-H) algorithm is used to gain samples from the posterior distribution for the auxiliary variables; it involves constructing a markov chain with the sought posterior as its stationary distribution, and sampling sequentially from this constructed chain to get an approximate posterior sample. The M-H algorithm is outlined in Algorithm 1.%Algorithm~\ref{alg:MH}. \willie{am i citing wrong?}

\begin{algorithm}[h!]
\label{alg:MH}
\caption{Metropolis Hastings Algorithm}
\begin{algorithmic}[1]
\STATE $f(x)$ is the stationary PDF from which samples are desired \hfill $\triangleright$ eq. \eqref{stationary_pdf}
\STATE $q( \cdot | x_{i-1})$ is the proposal pdf \hfill $\triangleright$ eq. \eqref{proposal_distro}
\STATE $x_{0} \leftarrow$ initialize
\STATE B $\leftarrow$ \# of burn in samples
\FOR{$i = 1 : \text{B} + N$}
\STATE Sample $x^{*} \sim q( \cdot | x_{i-1})$
\STATE Sample $\alpha \sim \text{Uniform}(0,1)$
\STATE $r_{accept} \leftarrow \frac{f(x^{*})q(x_{i-1} | x^{*})} {f(x_{i-1})q(x^{*} | x_{i-1})} $ \hfill $\triangleright$  eq. \eqref{accept_ratio}
\IF{$\alpha < \text{min} \{ 1, r_{accept} \} $}
\STATE $x_{i} \leftarrow x^{*}$
\ELSE
\STATE $x_{i} \leftarrow x_{i-1}$
\ENDIF
\ENDFOR
\STATE Discard samples $x_{1}, \ldots, x_{\text{B}}$
\STATE \textbf{Output}: Remaining samples $x_{1}, \ldots, x_{N}$
\end{algorithmic}
%\caption{Say thing at end}
\end{algorithm}

The M-H algorithm is used to sample auxiliary variables for each alive cluster. In short, we sample a new value for all $M$ auxiliary variables, denoted $z_{k,t,m}^{*}$, according to the following proposal distribution
\begin{equation}
\label{proposal_distro}
z_{k,t,m}^{*}  \sim  P(z_{k,t,m} | \theta_{k,t}) = \mathcal{N}(z_{k,t,m}^{s} | \boldsymbol{\mu}_{k,t}, \Sigma_{k,t}) \mathcal{M}n(z_{k,t,m}^{c} | \bold{p}_{k,t})
\end{equation}
compute the standard M-H acceptance ratio, which in this case (due to the proposal distribution choice) simplifies to 
\begin{equation}
\label{accept_ratio}
r_{accept} = \frac{P(\theta_{k,t-1} | z_{k,t,m}^{*})}{P(\theta_{k,t-1} | z_{k,t,m})}
\end{equation}
and keep this sample with probability $\text{min}(1, r_{accept})$.

@@

This section details a Sequential Monte Carlo (SMC) sampler---also known as a particle filter---used to perform inference, which provides another method to carry out unsupervised detection and tracking. The SMC inference algorithm is given in Algorithm~\ref{alg:SMC}. In this algorithm, at each time step $t \in \{ 1, \ldots, T \}$, a number of samples refered to as ``particles'' are generated; each particle consists of a sample from the posterior distribution over the assignment for each observation, $c_{1,t}, \ldots, c_{N_{t}, t}$, parameters for each cluster $\theta_{1,t}, \ldots, \theta_{K_{t}, t}$, and size after deletion for each cluster $m_{1,t}, \ldots, m_{K_{t},t}$. A set of particles is sampled at each time step from relevant proposal distributions (described in Sections~\ref{sec:smc_proposal_1}, \ref{sec:smc_proposal_2}, and \ref{sec:smc_proposal_3}), a weight is computed for each particle, and a subset of the particles are resampled and kept via a resampling process. Within each time step, Gibbs sampling is used to generate the samples associated with each particle.

The sequence of target distributions for the SMC algorithm may be written as
\begin{equation}
\begin{split}
%\pi_{t}(c_{1:N_{1}, 1}, \ldots, c_{1:N_{t}, t}, & \theta_{1:K_{1}, 1}, \ldots, \theta_{1:K_{t}, t}, m_{1:K_{1}, 1}, \ldots, m_{1:K_{t}, t}) = \\
%\pi_{t-1}(c_{1:N_{1}, 1}, \ldots, c_{1:N_{t-1}, t-1}, & \theta_{1:K_{1}, 1}, \ldots, \theta_{1:K_{t-1}, t-1}, m_{1:K_{1}, 1}, \ldots, m_{1:K_{t-1}, t-1}) \\
\pi_{t}(\bold{c}_{1:t}, \Theta_{1:t}, \bold{m}_{1:t}) & =  \pi_{t-1}(\bold{c}_{1:t-1}, \Theta_{1:t-1}, \bold{m}_{1:t-1})\\
& \times \prod_{i=1}^{N_{t}} P(c_{i,t} | \bold{m}_{t}, \Theta_{t}, \bold{c}_{1:t}, \bold{x}_{1:N_{t}}) \\
& \times \prod_{k=1}^{K} 
\begin{cases}
\begin{split}
P(\theta_{k,t} | \theta_{k,t-1}) \hspace{2mm} & \text{if} \hspace{2mm} k \leq K_{t-1} \\
\mathbb{G}_{0} \hspace{18mm} & \text{if} \hspace{2mm} k > K_{t-1}
\end{split}
\end{cases} \\
& \times \prod_{k=1}^{K} \text{DEL}(m_{k,t} | m_{k,t-1}, c_{1:N_{t-1}, t-1}, \rho)
\end{split}
\end{equation}
where $\bold{c}_{t} = c_{1,t}, \ldots, c_{N_{t}, t}$, $\Theta_{t} = \theta_{1,t}, \ldots, \theta_{K_{t}, t}$, and $\bold{m}_{t} = m_{1,t}, \ldots, m_{K_{t}, t}$ A few distributions referenced in the SMC algorithm are described in detail in the following sections.

@@

The probability of assignments given current cluster sizes, cluster parameters, and the Dirichlet Process granularity parameter $\alpha$ can be written as
\begin{equation}
\label{smc_assig_eqn}
\begin{split}
P \left( c_{i,t} | m_{1:K_{t}, t}, \theta_{1:K_{t}, t}, \alpha \right) &\propto
\text{CRP}(m_{1:K_{t}, t}, \alpha) \\ &\times
\begin{cases}
P(\bold{x}_{i,t} | \theta_{c_{i,t},t}) \hspace{2mm} \text{if} \hspace{2mm} k \leq K_{t-1} \\
\int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta)d\theta \hspace{2mm} k > K_{t-1}
\end{cases}
\end{split}
\end{equation}
The integral can $\int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta)d\theta$ can be determined analytically, and is given in \eqref{marginal_obs_prob}.

@@

\section*{Discussion and Conclusion}
\label{sec:discussion}

@@

%\section*{Conclusion}
